\papertitle

### 2014-03-20 final-report/paper.tex
Tackled the background

In particular, \system is built to inspect query execution in
Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data
management system currently being developed in the UW CSE database group. Myria
aims towards building a distributed database platform to provide \emph{big data
management and analytics as a service} primarily for scientific applications.
The proposed visualizations can easily be applied to other DDBMS as well (e.g.
Spark, Hadoop).

@@

Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues,
saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
of query execution.
\todo{it's not a dag, as we noticed}

@@

Twitter developed Ambrose\cite{ambrose}, a platform for visualization and real-time monitoring of MapReduce data workflows. They offer three different views to show associated jobs, job dependencies and progress. Ambrose has been released as open source. Ambrose, however, is not suitable for our needs as the abstraction level of jobs is too high and does not capture single operators.

Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing infrastructure and offers fine grained tracing of calls in Google's distributed systems. They also proposed an interface for visualizing traces. Similarly, X-trace\cite{fonseca2007x} was developed as a framework to trace which events cause what other events in a distributed environment. recently, there has been work on visualizing event traces collected in X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}. Due to the importance of these kinds of debug facilities, Twitter closely modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open source.

Dapper and X-trace focus on how data flows through a distributed system. In \system the visualization focuses on the operators and how the data flows through them. This orthogonal view is better suited for debugging performance bottlenecks in DDBMS and also scales to a larger number of events. Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data. Also, \system is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.

Tools to visualize query plans for example in SQL Server, as used to improve performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing queries and not query execution and have no visualization of data flow, which is necessary to optimize physical query execution.

@@

\system's architecture (Figure~\ref{fig:arch}) is composed of: (1) a back-end
used as a plug-in interface for the targeted distributed database system, and (2) a
front-end that produces performance visualizations what we describe in Section~\ref{sec:front}. The front-end and
back-end of \system are highly decoupled so that the visualization can be used not only with
Myria but also with other systems such as Spark or Hadoop. In
Section~\ref{sec:back} we illustrate how we implemented log collections and
aggregation in Myria.

@@

The role of the back-end is produce event logs that record when operators are called, when the call returns and when data is sent to another worker. These logs are then used to create the visualizations in \system's web UI as described in Section~\ref{sec:front}.

There are two possible approaches that we explored to collect event logs. The first and simplest approach was to write the logs on files directly to disk. The files would then get hauled back on one node using remote copy. The logs would then be parsed, and the data transformed into the desired format.

We switched to a more scalable data collection approach approach as shown in Figure~\ref{fig:arch} where logs get written directly into the worker's database as tuples in a relation. When the front-end client requests data, the master executes a query that collects the data and streams the results directly to the requesting client. We found this data collection approach to be more scalable, reliable and faster. More particularly, we are able to write queries to aggregate, filter and transform the collected data.

For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragments}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state. The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.

@@

\subsection{Front-end}
\label{sec:front}

The Myria web front-end server is written in Python and runs on Google App Engine\footnote{\url{https://developers.google.com/appengine/}}. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user. D3 is a JavaScript framework for data visualization on the web. We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that were made possible using D3.

\system's front-end gets data from the Myria server using \system's back-end as JSON and CSV files. JSON objects are used to describe the physical queryplan. CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.

>>>>

In particular, \system is built to inspect query execution in
Myria~\cite{myria}, a distributed big data
management system currently being developed in the UW CSE database group. Myria
aims towards building a distributed database platform to provide \emph{big data
management and analytics as a service} primarily for scientific applications.
The proposed visualizations can easily be applied to other DDBMS as well (e.g.
Spark, Hadoop).

@@

Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues,
saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
of query execution.
\todo{it's not a dag, as we noticed}

@@

Ambrose\cite{ambrose} is a platform developed by Twitter to visualize and
real-time monitor MapReduce data workflows. Ambrose
offers three different views to show associated jobs, job dependencies and
progress, which are not suitable for our needs as the abstraction level of jobs
is too high and does not capture single operators.

Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing
infrastructure that offers fine-grained tracing of calls in Google's distributed
systems. They also proposed an interface for visualizing traces. Similarly,
X-trace\cite{fonseca2007x} was developed as a framework to trace which events
cause what other events in a distributed environment. Recently, there has been
work on visualizing event traces collected in
X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
Due to the importance of these kinds of debug facilities, Twitter closely
modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open
source.

Dapper and X-trace focus on how data flows through a distributed system. In
contrast, in \system the visualization focuses on the operators and how the
data flows through them. This orthogonal view is better suited for debugging
performance bottlenecks in DDBMSs and also scales to a larger number of events.
Furthermore, \system offers different abstraction levels, which enables users
to find problems faster and handle larger amounts of profiling data. Also,
\system is specifically designed to help developers understand query execution
in distributed database system as opposed to general traces in distributed
systems.

Tools to visualize query plans, for example those used to improve the
performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing
queries and not query execution and have no visualization of data flow, which
is necessary to optimize physical query execution.

@@

\system's architecture (Figure~\ref{fig:arch}) consists of a back-end,
used as a plug-in interface for the distributed database system targeted, and a
front-end which implements the views described above. The front-end and
back-end of \system are highly decoupled so that the visualization can be used not only with
Myria but also with other systems such as Spark or Hadoop. In
Section~\ref{sec:back} we illustrate how we implemented log collections and
aggregation in Myria.

@@

The task of the back-end is to log when operators are called, when the call returns and when data is sent to another worker. Using this data and simple transformations, we can obtain the information necessary to create the visualizations to answer the questions in Section~\ref{sec:approach}.

There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization. First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy. Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.

We switched to a different approach as shown in Figure~\ref{fig:arch} where logs are written directly into the worker's database. When the front-end requests the data, the master executes a query that collects the data and streams the results directly to the requesting client. Logs are thus tuples in a relation and no longer separate in files and in a custom format. Despite being more reliable, faster and using the same logic that Myria already implements, we are able to write queries to aggregate, filter and transform the collected data.

For example, the data for how many workers are working on a certain fragment as described in Section~\ref{sec:fragments} is generated from the event log using a query to select only events for the root operators (the operators that have no parent in the same fragment) in each fragment and a custom map function that carries state. The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.

@@

\subsection{Front-end}

The Myria web front-end server is written in Python and runs on Google App Engine\footnote{\url{https://developers.google.com/appengine/}}. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user. D3 is a JavaScript framework for data visualization on the web. We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.

\system's front-end gets data from the Myria server as JSON and CSV files. JSON objects are used to describe the physical queryplan. CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.
